---
title: "LUGA expo: urbanisation and biodiversity"
author: "Julian Wittische"
format: 
  html:
    embed-resources: true
editor: visual
execute:
  echo: false
  warning: false
  message: false
  results: hide
---

## Maps

Biodiversity data is from end of 2024 mdata. Imperviousness data is from Copernicus from 2018.

```{r setup, include=FALSE}
source("config.R")
source("functions.R")
source("0_Libraries.R")
load(paste0(DATA_PATH, "/_Julian/LUGA-expo-DATA/Metrics.RData"))
```

```{r}
#| label: fig-imp
#| fig-cap: "Imperviousness percentage in Luxembourg in 2018."
mapviz(all, IMP)
```

::: {style="height: 3cm;"}
:::

```{r}
#| label: fig-all-obs
#| fig-cap: "Number of observations (all taxa)."

mapviz(all, observations)
```

::: {style="height: 3cm;"}
:::

```{r}
#| label: fig-all-spe
#| fig-cap: "Number of species (all taxa)."
mapviz(all, species)
```

::: {style="height: 3cm;"}
:::

```{r}
#| label: fig-all-spc
#| fig-cap: "Number of species corrected by the number of observations (all taxa)."
mapviz(all, speciesPobs)
```

::: {style="height: 3cm;"}
:::

```{r}
#| label: fig-birds-obs
#| fig-cap: "Number of observations (birds)."
mapviz(birds, observations)
```

::: {style="height: 3cm;"}
:::

```{r}
#| label: fig-birds-spe
#| fig-cap: "Number of species (birds)."
mapviz(birds, species)
```

::: {style="height: 3cm;"}
:::

```{r}
#| label: fig-birds-spc
#| fig-cap: "Number of species corrected by the number of observations (birds)."
mapviz(birds, speciesPobs)
```

::: {style="height: 3cm;"}
:::

```{r}
#| label: fig-vasc-obs
#| fig-cap: "Number of observations (vascular plants)."
mapviz(vasc, observations)
```

::: {style="height: 3cm;"}
:::

```{r}
#| label: fig-vasc-spe
#| fig-cap: "Number of species (vascular plants)."
mapviz(vasc, species)
```

::: {style="height: 3cm;"}
:::

```{r}
#| label: fig-vasc-spc
#| fig-cap: "Number of species corrected by the number of observations (vascular plants)."
mapviz(vasc, speciesPobs)
```

::: {style="height: 3cm;"}
:::

```{r}
#| label: fig-inver-obs
#| fig-cap: "Number of observations (invertebrates)."
mapviz(inver, observations)
```

::: {style="height: 3cm;"}
:::

```{r}
#| label: fig-inver-spe
#| fig-cap: "Number of species (invertebrates)."
mapviz(inver, species)
```

::: {style="height: 3cm;"}
:::

```{r}
#| label: fig-inver-spc
#| fig-cap: "Number of species corrected by the number of observations (invertebrates)."
mapviz(inver, speciesPobs)
```

::: {style="height: 3cm;"}
:::

```{r}
#| label: fig-neo-obs
#| fig-cap: "Number of observations (neobiota)."
mapviz(neo, observations)
```

::: {style="height: 3cm;"}
:::

```{r}
#| label: fig-neo-spe
#| fig-cap: "Number of species (neobiota)."
mapviz(neo, species)
```

::: {style="height: 3cm;"}
:::

```{r}
#| label: fig-neo-spc
#| fig-cap: "Number of species corrected by the number of observations (neobiota)."
mapviz(neo, speciesPobs)
```
::: {style="height: 3cm;"}
:::
::: {style="height: 3cm;"}
:::
## Model results

We are using the number of species corrected by the number of observations and relating this variable to imperviousness.

Two points for your own sanity:

1)  I chose this metric because the number of observations is driven by human demography, and the number of species is driven by the number of observations (although this saturates more or less quickly for a large range of spatial scales).

2)  For simplicity I used linear regression instead of loess, etc. However, model diagnostics show that were violating linear regression theoretical assumptions.

```{r}
#| label: fig-all-mod
#| fig-cap: "Linear relationship between imperviousness and the number of species corrected by the number of observations (all taxa)."
ggplot(all, aes(x = IMP, y = speciesPobs)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "blue")

```
::: {style="height: 3cm;"}
:::
```{r}
#| label: fig-birds-mod
#| fig-cap: "Linear relationship between imperviousness and the number of species corrected by the number of observations (birds)."
ggplot(birds, aes(x = IMP, y = speciesPobs)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "blue")
```
::: {style="height: 3cm;"}
:::
```{r}
#| label: fig-vasc-mod
#| fig-cap: "Linear relationship between imperviousness and the number of species corrected by the number of observations (vascular plants)."
ggplot(vasc, aes(x = IMP, y = speciesPobs)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "blue")
```
::: {style="height: 3cm;"}
:::
```{r}
#| label: fig-inver-mod
#| fig-cap: "Linear relationship between imperviousness and the number of species corrected by the number of observations (invertebrates)."
ggplot(inver, aes(x = IMP, y = speciesPobs)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "blue")
```
::: {style="height: 3cm;"}
:::
```{r}
#| label: fig-neo-mod
#| fig-cap: "Linear relationship between imperviousness and the number of species corrected by the number of observations (neobiota)."
ggplot(neo, aes(x = IMP, y = speciesPobs)) + 
  geom_point() +
  stat_smooth(method = "lm", col = "blue")

```
